### 2022-09-23 manuscript/manuscript.Rmd
update figure titles and add fig 3

---
title: "S2 pioneer factor MS"
author: "Tyler Gibson"
date: "8/24/2021"
output: html_document
---

# Abstract

# Introduction

# Results

## Ectopic pioneer factors bind and open closed chromatin

## Chromatin opening correlates with transcriptional activation

## Twist binds closed chromatin extensively and opens chromatin at some sites

## Motif content distinguishes chromatin opening by Twist from Zelda and Grainy head

## Ectopic transcription factors bind nonspecifically to active chromatin

# Discussion

# Materials and Methods

## Reagents

| Reagent                            | Source/vendor            | Catalog/accession \# |
|------------------------------------|--------------------------|----------------------|
| pMT-puro                           | AddGene                  | 17923                |
| Gibson Cloning                     | New England Biolabs      | E2611                |
| Schneider's Medium                 | Thermo Fisher Scientific | 21720001             |
| Fetal Bovine Serum                 | Omega Scientific         | FB-02                |
| Antibiotic/antimycotic             | Thermo Fisher Scientific | 15240062             |
| Effectene Transfection Reagent     | Qiagen                   | 301427               |
| PVDF Membrane                      |                          |                      |
| **Antibodies**                     |                          |                      |
| Anti-Zld                           | Harrison et al., 2010    |                      |
| Anti-Grh                           | Harrison et al., 2010    |                      |
| Anti-Twi                           | He et al., 2015          |                      |
| Anti-HA-peroxidase clone 3F10      | Roche                    | 12013819001          |
| Anti-Tubulin DM1A                  | Sigma                    | T6199-200UL          |
| goat anti-rabbit IgG-HRP conjugate | Bio-Rad                  | 1706515              |
| Anti-HA clone 12CA5                |                          |                      |
|                                    |                          |                      |
|                                    |                          |                      |
|                                    |                          |                      |

## Cloning

*zld*, *grh*, or *twi* cDNA (isoforms RB, RH, and RA, respectively) were cloned into pMT-puro plasmid (Addgene) via Gibson cloning (New England Biolabs). For cloning of the Zld and Grh DNA-binding domains, DNA encoding amino acids 1114-1487 (Zld) or amino acids 603-1032 (Grh) was cloned into pMT-puro.

-   **TG2A- Zld RB**

-   **TG3A- Grh PH**

-   **TG4A**

-   **TG6A**

-   **TG8A**

-   **MF8A**

-   GFP-DBD lines

## Cell culture and generation of stable cell lines

Drosophila S2 cells were cultured at 27°C in Schneider's medium (Thermo Fisher Scientific) supplemented with 10% fetal bovine serum (Omega Scientific) and antibiotic/antimycotic (Thermo Fisher Scientific). For the generation of stable cell lines, cells were plated at 5✕10^5^ cells per mL. After 24 hours, cells were transfected with 10 µg plasmid DNA using Effectene transfection reagent (Qiagen). After an additional 24 hours, puromycin was added to a final concentration of 2 µg/mL. Stable cell lines were recovered after 2-3 weeks of selection. Following recovery of stable cell lines, cells were cultured with 1 µg/mL puromycin.

## Induction of transcription factor expression

## Immunoblotting

Proteins were separated on denaturing polyacramide gels before transfer to a 0.45 μm Immobilon-P PVDF membrane (Millipore) in transfer buffer (25 mM Tris, 200 mM Glycine, 20% methanol) for 60 min (75 min for Zld) at 500mA at 4°C. The membranes were blocked with BLOTTO (2.5% non-fat dry milk, 0.5% BSA, 0.5% NP-40, in TBST) for 30 min at room temperature and then incubated with anti-Zld (1:750) (Harrison et al., 2010), anti-Grh (1:1000) (Harrison et al., 2010), anti-Twi (1:1000) (He et al. 2015), anti-HA-peroxidase (1:500) (clone 3F10, Roche), or anti-Tubulin (DM1A, 1:5000) (Sigma), overnight at 4°C. The secondary incubation was performed with goat anti-rabbit IgG-HRP conjugate (1:3000) (Bio-Rad) for 1 hr at room temperature. Blots were treated with SuperSignal West Pico PLUS chemiluminescent substrate (Thermo Fisher Scientific) and visualized using the Azure Biosystems c600 or Kodak/Carestream BioMax Film (VWR).

## Immunostaining

## ChIP-seq

## ChIP-seq analysis

## RNA-seq

## RNA-seq analysis

## ATAC-seq

## ATAC-seq analysis

## CUT&RUN

## CUT&RUN analysis

## Treatment of S2 cells with small molecule inhibitors

## Analysis of previously published data

## Data and code availability

# Figure Legends

**Figure 1: Ectopic pioneer factors open chromatin and activate transcription.**

**Figure 2: Twist binds closed chromatin extensively and drives accessibility at a limited number of sites.**

**Figure 3: Motif content shapes pioneer-factor activity**

**Figure 4: H3K27me3 and H3K9me3 are barriers to pioneer-factor binding.**

**Figure 5: Binding and opening of closed chromatin is concentration dependent.**

**Figure 6: Regions outside the DNA-binding domain are required for pioneer-factor function.**

**Figure 7: Pioneer-factor function requires stable chromatin occupancy.**