### 2023-01-06 manuscript/manuscript.qmd
break manuscript up into multiple qmd files

---
title: "Defining properties regulating pioneer-factor binding and function"
format: html
editor: visual
bibliography: references.bib
#csl: nature.csl
---

# Abstract
{{< include manuscript_sections/_01_Abstract.qmd >}}

# Introduction
{{< include manuscript_sections/_02_Introduction.qmd >}}

# Results
{{< include manuscript_sections/_03_Results.qmd >}}

# Discussion
{{< include manuscript_sections/_04_Discussion.qmd >}}

# Materials and Methods
{{< include manuscript_sections/_05_Discussion.qmd >}}

# Figure Legends