### 2023-01-06 manuscript/manuscript_sections/_06_Figure_Legends.qmd

**Figure 1: Ectopic pioneer factors open chromatin and activate transcription. a-b,** Genome browser tracks showing examples of individual class I, II or III regions for Zld **(a)** or Grh **(b)**.
ChIP seq and ATAC-seq signal are shown for Zld- or Grh-expressing cells along with ATAC-seq signal for WT cells lacking Zld or Grh expression.
**c,f,** Pie charts showing the distribution of class I, II, and III binding sites for Zld **(c)** or Grh **(f)**.
**d,g,** The proportion of class I, II or III binding sites that occur at promoters (-500 to +100 bp around transcription start site) or promoter distal regions for Zld **(d)** or Grh **(g)**.
**e,h,** Heatmaps showing ChIP-seq and ATAC-seq signal at class II and III regions for Zld **(e)** or Grh **(h). i,k,** Example genome browser tracks showing ChIP-seq, ATAC-seq and RNA-seq signal at class III regions where changes in chromatin accessibility are associated with changes in gene expression for Zld **(i)** or Grh **(k)**.
**j,i,** Proportion of class I, II or III regions that are proximal to a gene that is upregulated upon Zld **(j)** or Grh **(i)** expression.
Promoter-proximal and promoter-distal binding sites are shown separately.
The grey dotted line indicates the percentage of all binding sites that are proximal to a upregulated expressed gene.

**Figure 2: Twist binds closed chromatin extensively and drives accessibility at a limited number of sites. a,** Genome browser tracks showing examples of individual class I, II or III regions **b,** Pie charts showing the distribution of class I, II, and III binding sites.
**c,** The proportion of class I, II or III binding sites that occur at promoters (-500 to +100 bp around transcription start site) or promoter distal regions.
**d,** Heatmap showing ChIP-seq and ATAC-seq signal at class II and III regions.

**Figure 3: Motif content shapes pioneer-factor activity. a,** Venn diagram showing overlap between class I regions for Zld, Grh and Twi.
**b,** Heatmap of ChIP-seq signal at class I sites that overlap between Zld, Grh, and Twi.
Zld, Grh, and Twi ChIP-seq data were generated in this study.
Remaining ChIP-seq datasets were previously published (see Supplementary table XX).
**c-e,** Heatmaps showing the percentage of regions containing a canonical Zld **(c)**, Grh **(d)**, or Twi motif **(e)**.
Logos plot of the canonical motif are shown alongside the heatmap.
Percentages are shown for class I, II or III binding sites, or all WT ATAC-seq peaks as a control for the background motif frequency.
**f,h,j,** Boxplots of the average number of motifs per peak in class I, II, or III regions for Zld **(f)**, Grh **(h)**, or Twi **(j)**.
**g,i,k,** Boxplots of the average motif log-odds score for class I, II or III regions for Zld **(g)**, Grh **(i)**, or Twi **(k)**.
Motif scores were determined by comparing each identified motif instance to the canonical motif.

**Figure 4: Many endogenous binding sites are resistant to ectopic pioneer-factor binding. a-c,** Heatmaps comparing ChIP-seq signal in across different tissues for Zld **(a)**, Grh **(b)**, or Twi **(c)**.
ChIP-seq data for repressive histone modifications H3K27me3 and H3K9me3 is also shown.
**d-f,** Heatmaps comparing DMSO or tazemetostat (taz)-treated cells for Zld **(d)**, Grh **(e)**, or Twi **(f)**.
Zld, Grh, and Twi ChIP-seq data are z-score normalized.
CUT&RUN data for H3K27me3 is spike-in normalized using barcoded nucleosomes (see methods).

**Figure 5: Binding and opening of closed chromatin is concentration dependent. a,e,i,** Metaplots of ChIP-seq signal at class I-VI regions for Zld **(a)**, Grh **(e)**, or Twi **(i)** when expressed at different levels.
Concetrations of CuSO~4~ used to induce protein expression are indicated above each plot.
**b,f,j,** Metaplots of ATAC-seq signal at class I-VI regions for Zld **(b)**, Grh **(f)**, or Twi **(j)** when expressed at different levels.
**c,g,k,** Distribution of class I, II and III peaks when defined using ChIP-seq and ATAC-seq data for each concentration of CuSO~4~ for Zld **(c)**, Grh **(g)**, or Twi **(k)**.
Y-axis indicates the total number of peaks called at each concentration.
Colors indicate the proportion of those peaks that are defined as class I, II, or III.
**d,h,l,** Venn diagrams showing overlap of peaks called at physiological or supraphysiological expression levels for Zld **(d)**, Grh **(h)**, or Twi **(l)**.

**Figure 6: Regions outside the DNA-binding domain are required for pioneer-factor function. a-b,** Genome browser tracks showing example of class III regions where expression of the DBD alone is insufficient for robust binding and chromatin opening.
Shown for Zld **(a)** and Grh **(b)**.
**c,g,** Metaplots comparing ChIP-seq signal for full-length protein or DBD alone at class I, II and III regions for Zld **(c)** or Grh **(g)**.
**d,f,** Volcano plots for ATAC-seq data showing the absence of changes to chromatin accessibility upon expression of Zld **(d)** or Grh DBD **(f)**.

**Figure 7: Pioneer-factor function requires stable chromatin occupancy.**

**Extended Data Figure 1: Stable cell lines allow inducible expression of transcription factors at physiological concentrations.**

-   ~~Schematic of cell line generation.~~

-   ~~Western blots to determine physiological expression and lack of protein expression in WT cells.~~

-   ~~RNA-seq data to show leaky expression and strong induction of RNA levels.~~

**Extended Data Figure 2: Zld and Grh bind to chromatin rapidly after induction of protein expression**

-   Western blot time course of protein levels

-   CUT&RUN time course to show rapid binding

-   likely rules out dependence on DNA replication

**Extended Data Figure 3: Ectopic expression of pioneer factors in S2 cells leads to widespread changes to chromatin accessibility and gene expression.**

-   volcano and/or MA plots with enrichment of binding

-   Analysis of nonspecific changes?

-   RNA-seq volcano/MA plots

-   RNA-seq GO analysis

**Extended Data Figure 4: Twist binding leads to chromatin opening and transcriptional activation**

-   Twi volcano and examples

-   TWI RNA-seq volcano and examples

-   Twi RNA-seq GO analysis

**Extended Data Figure 5: Chromatin features associated with Zld, Grh and Twi binding sites**

-   analysis of published ChIP datasets not shown in main figure

-   comparison of class I-III sites to show that class II-III sites are devoid of most histone mods

-   Class III sites have higher levels of H1 and H3 than class II sites.

**Extended Data Figure 6: Zld, Grh and Twi do not bind preferentially to motifs in a particular position on nucleosomes**

-   De novo motif enrichment for each factor

-   MNase data to show enrichment of nucleosomes around target motifs

-   Clustering of MNase data to show that there is no preferred nucleosome position relative to motifs

-   additional analysis of Fiber-seq data to show that nucleosomes at these regions are poorly positioned

**Additional information to include in supplemental figures:**

-   De novo motif results

-   Analysis of promoter motifs

-   New binding sites at higher expression levels

-   Zld DBD ChIP only at CAGGTAG sites