### 2023-01-12 manuscript/manuscript.qmd
add code for supp fig6

---
title: "Defining properties regulating pioneer-factor binding and function"
format: html
editor: visual
bibliography: references.bib
#csl: nature.csl
---

# Abstract

{{< include manuscript_sections/_01_Abstract.qmd >}}

# Introduction

{{< include manuscript_sections/_02_Introduction.qmd >}}

# Results

{{< include manuscript_sections/_03_Results.qmd >}}

# Discussion

{{< include manuscript_sections/_04_Discussion.qmd >}}

# Materials and Methods

{{< include manuscript_sections/_05_Discussion.qmd >}}

# Figure Legends

{{< include manuscript_sections/_06_Figure_Legends.qmd >}}