### 2023-03-02 manuscript/manuscript_sections/_01_Abstract.qmd
update abstract draft

During development, transcription factors establish cell-type specific gene expression programs.
While many transcription factors cannot bind to target DNA in the context of nucleosomes, pioneer transcription factors access nucleosomal targets and promote chromatin opening.
Despite the ability to bind to nucleosomal target motifs, many pioneer factors display cell-type specific binding and activity.
It is unclear how pioneer factor occupancy is regulated in different cell types during development and how chromatin occupancy relates to chromatin opening activity.
To determine how pioneer-factor binding and activity are regulated, we systematically studied three Drosophila transcription factors with distinct DNA-binding domains and biological functions: Zelda, Grainy head, and Twist.
We demonstrated that the level of chromatin occupancy is a key determinant of pioneer-factor activity.
Zld, Grh and Twi occupancy are regulated by a numerous factors including motif content at binding sites, local chromatin context, and protein concentration.
We found that regions outside the DNA-binding domain are required for pioneer-factor binding and chromatin opening.
Together, our results demonstrate that both protein-intrinsic and cell-type specific factors regulate pioneer-factor occupancy and activity.
These findings have important implications for how pioneer factors drive development and how misexpression can lead to disease.