### 2023-03-06 manuscript/manuscript.qmd
update MS draft

---
title: "Protein-intrinsic properties and epigenetic effects regulate pioneer-factor binding and function"
author:
  - Tyler J. Gibson, Melissa M. Harrison
  # - Melissa M. Harrison
affiliations:
  - name: Department of Biomolecular Chemistry, University of Wisconsin School of Medicine and Public Health
    city: Madison
    state: WI
format:
  docx:
    reference-doc: custom-reference-doc.docx
editor: visual
bibliography: references.bib
csl: nsmb.csl
---

Department of Biomolecular Chemistry, University of Wisconsin School of Medicine and Public Health, Madison, WI

{{< pagebreak >}}

# Abstract
{{< include manuscript_sections/_01_Abstract.qmd >}}

{{< pagebreak >}}

# Introduction

{{< include manuscript_sections/_02_Introduction.qmd >}}

# Results

{{< include manuscript_sections/_03_Results.qmd >}}

# Discussion

{{< include manuscript_sections/_04_Discussion.qmd >}}

{{< pagebreak >}}

# Materials and Methods

{{< include manuscript_sections/_05_Methods.qmd >}}

{{< pagebreak >}}

# References

::: {#refs}
:::

{{< pagebreak >}}

# Figure Legends

{{< include manuscript_sections/_06_Figure_Legends.qmd >}}