### 2023-03-06 manuscript/manuscript_sections/_01_Abstract.qmd

Transcription factors establish cell-type specific gene expression programs to drive developmental transitions.
While chromatin presents a barrier to the binding of many transcription factors, pioneer factors access nucleosomal targets and promote chromatin opening.
Despite the ability to bind to target motifs in closed chromatin, many pioneer factors display cell-type specific binding and activity.
The mechanisms that govern pioneer-factor occupancy in different cell types and the relationship between chromatin occupancy and chromatin opening remain unclear.
To determine how pioneer-factor binding and activity are regulated, we systematically studied three Drosophila transcription factors with distinct DNA-binding domains and biological functions: Zelda, Grainy head, and Twist.
We demonstrated that the level of chromatin occupancy is a key determinant of pioneer-factor activity.
Zelda, Grainy head and Twist occupancy is regulated by multiple factors, including motif content at binding sites, local chromatin context, and protein concentration.
We found that regions outside the DNA-binding domain are required for binding and chromatin opening.
Together, our results demonstrate that pioneering activity is not a binary feature intrinsic to a protein but occurs on a spectrum and can be regulated by a variety of protein-intrinsic and cell-type-specific features.
These findings have important implications for how pioneer factors drive development and how misexpression can lead to disease.