### 2023-03-10 manuscript/supplemental_figures.qmd
update figs

---
format: 
  pdf:
    include-in-header:
      text: |
        \usepackage[labelfont=bf,format=plain]{caption}
        \pagenumbering{gobble}
    mainfont: Arial
  docx:
    reference-doc: custom-reference-doc.docx
crossref:
  fig-title: '**Extended Data Figure**'
  title-delim: "**.**"

---

::: {#fig-1}
![](figures/extended_data_fig1.pdf)

**Stable cell lines allow inducible expression of transcription factors at physiological concentrations. a,** Schematic of generation of stable cell lines and induction of protein expression.
**b,** mRNA levels of Zld and Grh in S2 cells.
Top histogram shows the distribution of mRNA levels for all Drosophila genes.
Vertical dashed line indicates a log~2~ RPKM value of 0 as a threshold for considering a gene to be expressed.
**c-d,** Immunoblots showing titration of Zld **(c)** or Grh **(d)** protein levels in stable cell lines.
Two independently generated cell lines are shown and compared to 2-3H embryos.
:::

::: {#fig-2}
![](figures/extended_data_fig2.pdf)

**Zld and Grh bind to chromatin rapidly after induction of protein expression. a-b,** Immunoblots showing time course of Zld **(a)** or Grh **(b)** protein expression following induction of stable cell lines.
**c-d,** Metaplots showing average z-score normalized CUT&RUN signal at class I, II or III sites at different time points following induction.
:::

::: {#fig-3}
![](figures/extended_data_fig3.pdf)

**Ectopic expression of pioneer factors in S2 cells leads to widespread changes to chromatin accessibility and gene expression. a,d,** Volcano plots showing changes in ATAC-seq signal in cells expressing Zld **(a)** or Grh **(d)** when compared to WT cells treated with the same concentration of CuSO~4~.
**b,e,** RNA-seq volcano plots showing gene expression changes in cells expressing Zld **(a)** or Grh **(d)** when compared to WT cells treated with the same concentration of CuSO~4~.
**c,f,** Bar plots showing enrichment of gene ontology terms in genes significantly upregulated upon expression of Zld **(c)** or Grh **(f)**.
:::

::: {#fig-4}
![](figures/extended_data_fig4.pdf)

**Twist binding leads to chromatin opening and transcriptional activation. a,** Immunoblot showing titration of Twi or HA-Twi protein levels in stable cell lines.
Protein levels in stable cell lines are compared to 3-4 hour old embryos.
**b,** Volcano plots showing changes in ATAC-seq signal in cells expressing Twi when compared to WT cells treated with the same concentration of CuSO~4~.
**c,** RNA-seq volcano plots showing gene expression changes in cells expressing Twi when compared to WT cells treated with the same concentration of CuSO~4~.
**d,** Bar plots showing enrichment of gene ontology terms in genes significantly upregulated upon expression of Twi.
:::

::: {#fig-5}
![](figures/extended_data_fig5.pdf)

**Chromatin features associated with Zld, Grh and Twi binding sites. a,c,e,** Heatmaps showing the levels of different chromatin marks in class I, II or III regions for Zld (a), Grh (c) or Twi (e).
The color represents the average z-score normalized read depth across a 1 KB region surrounding the center of class I, II or III ChIP-seq peaks.
**b,d,f,** Example genome browser tracks for class III regions with high levels of H3K27me3 for Zld (b), Grh (d), or Twi (f).
:::

::: {#fig-6}
![](figures/extended_data_fig6.pdf)

**Zld, Grh and Twi do not bind preferentially to motifs in a particular position on nucleosomes. a-c,** Metaplots showing average MNase signal at motifs within class I, II, or III regions for Zld (a), Grh (b), or Twi (c).
**d-f,** Heatmaps showing MNase signal at motifs within class II and III regions for Zld (d), Grh (e), or Twi (f).
Rows are ordered based on hierarchical clustering to highlight the various patterns of MNase signal around motifs.
:::

::: {#fig-7}
![](figures/extended_data_fig7.pdf)

**Zld, Grh and Twi display cell-type specific binding. a-c,** Venn diagrams showing overlap between ChIP-seq peaks identified in different tissues for Zld (a), Grh (b) or Twi (c).
**d-l,** Genome browser tracks showing examples of class IV, V, and VI regions for Zld (d-f), Grh (g-i) or Twi (j-l).
For each example, the top tracks show H3K27me3 and H3K9me3 signal over a larger region.
Dashed gray lines indicate a zoomed-in region where Zld, Grh or Twi ChIP-seq signal is shown in S2 cells or in embryos.
:::

::: {#fig-8}
![](figures/extended_data_fig8.pdf)

**PRC2 inhibition is insufficient to promote Zld, Grh, and Twi binding to repressed targets. a,** Immunoblot showing H3K27me3 levels in two replicates of DMSO- or tazemetostat-treated cells.
Tubulin levels are shown as a loading control.
:::

::: {#fig-9}
![](figures/extended_data_fig9.pdf)

**Expression of Zld, Grh, or Twi at supraphysiological levels results in chromatin opening at a small number of novel binding sites. a-c,** Immunoblots showing Zld (a), Grh (b) or Twi (c) protein levels when stable cell lines are induced using different concentrations of CuSO~4~.
**d-f,** Bar plots showing the percentage of previously defined class I-VI binding sites that are bound by Zld (d), Grh (e), or Twi (f) when expressed at varying concentrations.
**g-i,** Bar plots showing the percentage of previously defined class I-VI binding sites that overlap an ATAC-seq peak when Zld (g), Grh (h), or Twi (i) are expressed at varying concentrations.
:::

::: {#fig-10}
![](figures/extended_data_fig10.pdf)

**Expression of Zld and Grh DNA-binding domains at protein levels comparable to the full-length proteins. a-b,** Immunoblots showing titration of protein levels for Zld (a) or Grh (b) DNA-binding domains to match expression of the full-length proteins.
DNA-binding domain protein levels are shown at a range of CuSO~4~ concentrations and compared to an equivalent number of cells expressing full-length Zld or Grh at approximately physiological levels.
**c-d,** Immunofluorescent microscopy images of stable cell lines expressing full-length protein or DBD only for Zld (c) or (Grh).
Stable cell lines are compared to WT cells.
**e,** Metaplots showing the average ChIP-seq signal for full-length Zld or the DNA-binding domain alone at the subset of previously identified class I, II or III binding sites that contain a canonical Zelda CAGGTAG motif.
:::