### 2023-03-15 manuscript/manuscript_sections/_01_Abstract.qmd
incorporate MMH comments from 2023-03-13

While chromatin presents a barrier to the binding of many transcription factors, pioneer factors access nucleosomal targets and promote chromatin opening.
Despite binding to target motifs in closed chromatin, many pioneer factors display cell-type specific binding and activity.
The mechanisms that govern pioneer-factor occupancy and the relationship between chromatin occupancy and opening remain unclear.
We studied three Drosophila transcription factors with distinct DNA-binding domains and biological functions: Zelda, Grainy head, and Twist.
We demonstrated that the level of chromatin occupancy is a key determinant of pioneer-factor activity.
Occupancy is regulated by multiple factors, including motif content at binding sites, local chromatin context, and protein concentration.
Regions outside the DNA-binding domain are required for pioneer-factor binding and chromatin opening.
Together, our results show that pioneering activity is not a binary feature intrinsic to a protein but occurs on a spectrum and can be regulated by a variety of protein-intrinsic and cell-type-specific features.