### 2023-03-15 manuscript/manuscript_sections/_06_Figure_Legends.qmd

**Figure 1: Ectopically expressed pioneer factors open chromatin and activate transcription. a-b,** Genome browser tracks showing examples of individual class I, II or III regions for Zld **(a)** or Grh **(b)**.
ChIP seq and ATAC-seq signal are shown for Zld- or Grh-expressing cells along with ATAC-seq signal for wild-type (WT) cells lacking Zld or Grh expression.
**c,f,** Pie charts showing the distribution of class I, II, and III binding sites for Zld **(c)** or Grh **(f)**.
**d,g,** The proportion of class I, II or III binding sites that occur at promoters (-500 to +100 bp around transcription start site) or promoter distal regions for Zld **(d)** or Grh **(g)**.
**e,h,** Heatmaps showing ChIP-seq and ATAC-seq signal at class II and III regions for Zld **(e)** or Grh **(h). i,k,** Example genome browser tracks showing ChIP-seq, ATAC-seq and RNA-seq signal at class III regions where changes in chromatin accessibility are associated with changes in gene expression for Zld **(i)** or Grh **(k)**.
**j,i,** Proportion of class I, II or III regions that is proximal to a gene that is upregulated upon Zld **(j)** or Grh **(i)** expression.
Promoter-proximal and promoter-distal binding sites are shown separately.
The grey dotted line indicates the percentage of all binding sites that are proximal to a upregulated expressed gene.

**Figure 2: Twist binds closed chromatin extensively and drives accessibility at a limited number of sites. a,** Genome browser tracks showing examples of individual class I, II or III regions **b,** Pie charts showing the distribution of class I, II, and III binding sites.
**c,** The proportion of class I, II or III binding sites that occur at promoters (-500 to +100 bp around transcription start site) or promoter distal regions.
**d,** Heatmap showing ChIP-seq and ATAC-seq signal at class II and III regions.

**Figure 3: Motif content shapes pioneer-factor activity. a,** Venn diagram showing overlap between class I regions for Zld, Grh and Twi.
**b,** Heatmap of ChIP-seq signal at class I sites that overlap between Zld, Grh, and Twi.
Zld, Grh, and Twi ChIP-seq data were generated in this study.
Remaining ChIP-seq datasets were previously published (see Supplementary table 1)[@Henriques2018;@Philip2015;@Tettey2019].
**c-e,** Heatmaps showing the percentage of regions containing a canonical Zld **(c)**, Grh **(d)**, or Twi motif **(e)**.
Logos plot of the canonical motif are shown alongside the heatmap.
Percentages are shown for class I, II or III binding sites, or all wild-type ATAC-seq peaks as a control for the background motif frequency.
**f,h,j,** Boxplots of the average number of motifs per peak in class I, II, or III regions for Zld **(f)**, Grh **(h)**, or Twi **(j)**.
**g,i,k,** Boxplots of the average motif log-odds score for class I, II or III regions for Zld **(g)**, Grh **(i)**, or Twi **(k)**.
Motif scores were determined by comparing each identified motif instance to the canonical motif.

**Figure 4: Many endogenous binding sites are resistant to ectopic pioneer-factor binding. a-c,** Heatmaps comparing ChIP-seq signal in across different tissues for Zld **(a)**, Grh **(b)**, or Twi **(c)**.
ChIP-seq data for repressive histone modifications H3K27me3 [@Jain2020] and H3K9me3[@Eastwood2021] in WT S2 cells is also shown.
**d-f,** Heatmaps comparing DMSO or tazemetostat (taz)-treated cells for Zld **(d)**, Grh **(e)**, or Twi **(f)**.
Zld, Grh, and Twi ChIP-seq data are z-score normalized.
CUT&RUN data for H3K27me3 is spike-in normalized using barcoded nucleosomes (see methods).

**Figure 5: Pioneer-factor binding and opening of closed chromatin is concentration dependent. a,e,i,** Metaplots of ChIP-seq signal at class I-VI regions for Zld **(a)**, Grh **(e)**, or Twi **(i)** when expressed at different levels.
Concetrations of CuSO~4~ used to induce protein expression are indicated above each plot.
**b,f,j,** Metaplots of ATAC-seq signal at class I-VI regions for Zld **(b)**, Grh **(f)**, or Twi **(j)** when expressed at different levels.
**c,g,k,** Distribution of class I, II and III peaks when defined using ChIP-seq and ATAC-seq data for each concentration of CuSO~4~ for Zld **(c)**, Grh **(g)**, or Twi **(k)**.
Y-axis indicates the total number of peaks called at each concentration.
Colors indicate the proportion of those peaks that are defined as class I, II, or III.
**d,h,l,** Venn diagrams showing overlap of peaks called at physiological or supraphysiological expression levels for Zld **(d)**, Grh **(h)**, or Twi **(l)**.

**Figure 6: Regions outside the DNA-binding domain are required for pioneer-factor function. a-b,** Genome browser tracks showing example of class III regions where expression of the DBD alone is insufficient for robust binding and chromatin opening.
Adjacent class I regions are shown where the DBD alone can bind.
Tracks shown for Zld **(a)** and Grh **(b)**.
**c,g,** Metaplots comparing ChIP-seq signal for full-length protein or DBD alone at class I, II and III regions for Zld **(c)** or Grh **(g)**.
**d,f,** Volcano plots for ATAC-seq data showing the absence of changes to chromatin accessibility upon expression of Zld **(d)** or Grh DBD **(f)**.
**g,** Model for the regulation of PFs.
A high level of chromatin occupancy is required for PFs to initiate chromatin opening.
Chromatin occupancy can be regulated by protein-intrinsic features such as motif affinity and protein domains, and by cell-intrinsic properties such as histone modifications, PF concentration, and cell-type specific cofactors (outlined in blue).

**Extended Data Figure 1: Stable cell lines allow inducible expression of transcription factors at physiological concentrations. a,** Schematic of generation of stable cell lines and induction of protein expression.
**b,** mRNA levels of Zld and Grh in S2 cells.
Top histogram shows the distribution of mRNA levels for all Drosophila genes.
Vertical dashed line indicates a log~2~ RPKM value of 0 as a threshold for considering a gene to be expressed.
**c-d,** Immunoblots showing titration of Zld **(c)** or Grh **(d)** protein levels in stable cell lines.
Two independently generated cell lines are shown and compared to 2-3H embryos.
Black arrowheads indicate Zld **(c)** or Grh **(d)**.
Gray arrowheads indicate background bands used to assess loading.
**e-f,** Heatmaps comparing Zld **(e)**, Grh **(f)**, or Twi **(g)** ChIP-seq signal to control experiments in which anti-Zld, anti-Grh, anti-Twi, or IgG antibodies were used to perform immunopreciptation in WT cells.

**Extended Data Figure 2: Zld and Grh bind to chromatin rapidly after induction of protein expression. a-b,** Immunoblots showing time course of Zld **(a)** or Grh **(b)** protein expression following induction of stable cell lines.
Black arrowheads indicate Zld **(a)** or Grh **(b)**.
Gray arrowheads indicate background bands used to assess loading.
**c-d,** Metaplots showing average z-score normalized CUT&RUN signal at class I, II or III sites at different time points following induction.

**Extended Data Figure 3: Ectopic expression of pioneer factors in S2 cells leads to widespread changes to chromatin accessibility and gene expression. a,d,** Volcano plots showing changes in ATAC-seq signal in cells expressing Zld **(a)** or Grh **(d)** when compared to WT cells treated with the same concentration of CuSO~4~.
**b,e,** RNA-seq volcano plots showing gene expression changes in cells expressing Zld **(b)** or Grh **(e)** when compared to WT cells treated with the same concentration of CuSO~4~.
**c,f,** Bar plots showing enrichment of gene ontology terms in genes significantly upregulated upon expression of Zld **(c)** or Grh **(f)**.

**Extended Data Figure 4: Twist binding leads to chromatin opening and transcriptional activation. a,** Immunoblot showing titration of Twi or HA-Twi protein levels in stable cell lines.
Protein levels in stable cell lines are compared to 3-4 hour old embryos.
**b,** Volcano plots showing changes in ATAC-seq signal in cells expressing Twi when compared to WT cells treated with the same concentration of CuSO~4~.
**c,** RNA-seq volcano plots showing gene expression changes in cells expressing Twi when compared to WT cells treated with the same concentration of CuSO~4~.
**d,** Bar plots showing enrichment of gene ontology terms in genes significantly upregulated upon expression of Twi.

**Extended Data Figure 5: Chromatin features associated with Zld, Grh and Twi binding sites. a,c,e,** Heatmaps showing the levels of different chromatin marks in class I, II or III regions for Zld **(a)**, Grh **(c)** or Twi **(e)**.
The color represents the average z-score normalized read depth across a 1 KB region surrounding the center of class I, II or III ChIP-seq peaks.
**b,d,f,** Example genome browser tracks for class III regions with high levels of H3K27me3 for Zld **(b)**, Grh **(d)**, or Twi **(f)**.

**Extended Data Figure 6: Zld, Grh and Twi do not bind preferentially to motifs in a particular position on nucleosomes. a-c,** Metaplots showing average MNase signal from WT cells centered on motifs within class I, II, or III regions for Zld **(a)**, Grh **(b)**, or Twi **(c)**.
**d-f,** Heatmaps showing MNase signal centered on motifs within class II and III regions for Zld **(d)**, Grh **(e)**, or Twi **(f)**.
Rows are ordered based on hierarchical clustering to highlight the various patterns of MNase signal around motifs.

**Extended Data Figure 7: Zld, Grh and Twi display cell-type specific binding. a-c,** Venn diagrams showing overlap between ChIP-seq peaks identified in different tissues for Zld **(a)**, Grh **(b)** or Twi **(c)**.
**d-l,** Genome browser tracks showing examples of class IV, V, and VI regions for Zld **(d-f)**, Grh **(g-i)** or Twi **(j-l)**.
For each example, the top tracks show H3K27me3 and H3K9me3 signal over a larger region.
Dashed gray lines indicate a zoomed-in region where Zld, Grh or Twi ChIP-seq signal is shown in S2 cells or in embryos.

**Extended Data Figure 8: Specific detection of H3K27me3 in tazemtostat-treated cells. a,** Immunoblot showing H3K27me3 levels in two replicates of DMSO- or tazemetostat-treated cells.
Tubulin levels are shown as a loading control.
**b,** Heatmap showing specificity of anti-H327me3 antibody in CUT&RUN reactions.
A panel of barcoded spike-in nucleosomes bearing different modifications was added to each CUT&RUN reaction (see methods).
For each sample, the heatmap displays the percentage of barcode reads for each sample and histone modification relative to the total number of barcode reads for all modifications.

**Extended Data Figure 9: Expression of Zld, Grh, or Twi at supraphysiological levels results in chromatin opening at a small number of novel binding sites. a-c,** Immunoblots showing Zld **(a)**, Grh **(b)** or Twi **(c)** protein levels when stable cell lines are induced using different concentrations of CuSO~4~.
**d-f,** Bar plots showing the percentage of previously defined class I-VI binding sites that are bound by Zld **(d)**, Grh **(e)**, or Twi **(f)** when expressed at varying concentrations.
**g-i,** Bar plots showing the percentage of previously defined class I-VI binding sites that overlap an ATAC-seq peak when Zld **(g)**, Grh **(h)**, or Twi **(i)** are expressed at varying concentrations.

**Extended Data Figure 10: Expression of Zld and Grh DNA-binding domains at protein levels comparable to the full-length proteins. a-b,** Immunoblots showing titration of protein levels for Zld **(a)** or Grh **(b)** DNA-binding domains to match expression of the full-length proteins.
DNA-binding domain protein levels are shown at a range of CuSO~4~ concentrations and compared to an equivalent number of cells expressing full-length Zld or Grh at approximately physiological levels.
**c-d,** Immunofluorescent microscopy images of stable cell lines expressing full-length protein or DBD only for Zld **(c)** or (Grh).
Stable cell lines are compared to WT cells.

**Supplementary table 1:** List of previously published genomics datasets that were analyzed in this work.