### 2023-03-20 manuscript/manuscript_sections/_01_Abstract.qmd
update MS

While chromatin presents a barrier to the binding of many transcription factors, pioneer factors access nucleosomal targets and promote chromatin opening.
Despite binding to target motifs in closed chromatin, many pioneer factors display cell-type specific binding and activity.
The mechanisms governing pioneer-factor occupancy and the relationship between chromatin occupancy and opening remain unclear.
We studied three *Drosophila* transcription factors with distinct DNA-binding domains and biological functions: Zelda, Grainy head, and Twist.
We demonstrated that the level of chromatin occupancy is a key determinant of activity.
Multiple factors regulate occupancy, including motif content, local chromatin, and protein concentration.
Regions outside the DNA-binding domain are required for binding and chromatin opening.
Our results show that pioneering activity is not a binary feature intrinsic to a protein but occurs on a spectrum and is regulated by a variety of protein-intrinsic and cell-type-specific features.