### 2023-10-20 manuscript/manuscript_sections/_05_Methods.qmd

---
# bibliography: ../references.bib
---

## Cell culture and generation of stable cell lines

*zld*, *grh*, or *twi* cDNA (isoforms RB, RH, and RA, respectively) were cloned into pMT-puro plasmid (Addgene) via Gibson cloning (New England Biolabs).
For cloning of the Zld and Grh DNA-binding domains, DNA encoding amino acids 1114-1487 (Zld) or amino acids 603-1032 (Grh) was cloned into pMT-puro.

We validated our wild-type S2 cells by performing RNA-seq and verifying the expression of known S2 cell markers and the absence of Kc cell markers [@Cherbas2011].
S2 cells were cultured at 27°C in Schneider's medium (Thermo Fisher Scientific) supplemented with 10% fetal bovine serum (Omega Scientific) and antibiotic/antimycotic (Thermo Fisher Scientific).
For the generation of stable cell lines, cells were plated at 5✕10^5^ cells per mL.
After 24 hours, cells were transfected with 10 µg plasmid DNA using Effectene transfection reagent (Qiagen).
After an additional 24 hours, puromycin was added to a final concentration of 2 µg/mL.
Stable cell lines were recovered after 2-3 weeks of selection.
Following recovery of stable cell lines, cells were cultured with 1 µg/mL puromycin.

## Induction of transcription factor expression

Transcription factor expression was induced by adding CuSO~4~ to the cell culture media.
Expression of protein at approximately physiological levels was achieved with 1000, 100, or 40 µM CuSO4 for Zld, Grh or Twi, respectively (Extended Data Fig. 1c-d; Extended Data Fig. 4a).
For experiments with multiple protein levels (Fig. 5; Extended Data Fig. 9a-c), the following CuSO~4~ concentrations were used: 500 µM (Zld low), 1000 µM (Zld medium) or 1500 µM (Zld high); 25 µM (Grh low), 100 µM (Grh medium) or 400 µM (Grh high); 10 µM (Twi low), 40 µM (Twi medium) or 160 µM (Twi high); 100 µM (Zld DNA-binding domain low); 400 µM (Zld DNA-binding domain high); 20 µM (Grh DNA-binding domain low); 100 µM (Grh DNA-binding domain high).
The Zld or Grh DNA-binding domains were induced with 100 µM, or 20 µM CuSO~4~, respectively for immunofluorescence experiments (Extended Data Fig. 10).
For all induction experiments, cells were plated at 1×10^6^ cells per mL.
Unless otherwise specified, experiments were performed with cells harvested 48 hours after induction.

## Immunoblotting

Proteins were separated on denaturing polyacramide gels before transfer to a 0.45 μm Immobilon-P PVDF membrane (Millipore) in transfer buffer (25 mM Tris, 200 mM Glycine, 20% methanol) for 60 min (75 min for Zld) at 500mA at 4°C.
The membranes were blocked with BLOTTO (2.5% non-fat dry milk, 0.5% BSA, 0.5% NP-40, in TBST) for 30 min at room temperature and then incubated with anti-Zld (1:750) [@Harrison2010], anti-Grh (1:1000) [@Harrison2010], anti-Twi (1:1000) [@He2015], anti-HA-peroxidase (1:500) (clone 3F10, Roche), or anti-Tubulin (DM1A, 1:5000) (Sigma), overnight at 4°C.
The secondary incubation was performed with goat anti-rabbit IgG-HRP conjugate (1:3000) or goat anti-mouse IgG-HRP conjugate (1:3000) (Bio-Rad) for 1 hr at room temperature.
Blots were treated with SuperSignal West Pico PLUS chemiluminescent substrate (Thermo Fisher Scientific) and visualized using the Azure Biosystems c600 or Kodak/Carestream BioMax Film (VWR).

## Immunostaining

Immunostaining was performed as described previously [@angshuman2007].
Briefly, 1×10^6^ cells in 200 µL PBS were applied to a coverslip pre-coated with 10% poly-L-lysine.
Cells were fixed with 4% methanol-free formaldehyde for 30 minutes at 37°C, permeabilized with 0.2% Triton-X 100 for 10 minutes at room temperature, and blocked with 1% BSA for 1 hour.
Cells were incubated with a 1:500 dilution of anti-Zld or anti-Grh primary antibody overnight at 4°C.
After washing with PBS, cells were incubated with a 1:500 dilution of goat anti-rabbit IgG DyLight 488 conjugated secondary antibody (Thermo Fisher Scientific #35552) for 1 hour at RT.
Imaging was performed on a Nikon Ti2 epi-fluorescent microscope with a 60x objective lens.

## ChIP-seq

For each ChIP experiment, \~50 million cells were crosslinked by adding methanol-free formaldehyde directly to the cell culture medium to a final concentration of 0.8%.
Cells were rotated on a nutator for 7 minutes at room temperature before quenching of crosslinking by adding glycine to 125 mM.
Crosslinked cells were pelleted by centrifugation at 600 × *g* for 3 minutes.
Cells were washed twice with 1X PBS.
Cells were lysed by resuspending in lysis buffer (50 mM HEPES pH 7.9, 140 mM NaCl, 1 mM EDTA, 10% glycerol, 0.5% NP40, 0.25% Triton-X 100) with protease inhibitors (Pierce) and incubating on ice for 10 minutes.
The resulting chromatin was pelleted for 5 minutes at 1500 × *g* and resuspended in RIPA buffer.
Chromatin was sonicated in a Covaris S220 Ultrasonicator (5 cycles of 120 seconds with 60 second delay, 170 peak power, 10% duty factor, 200 cycles/burst).
Sonicated chromatin was centrifuged for 10 minutes at 10,000 × *g* to pellet insoluble material.
5% of the supernatant was set aside as input, and the remainder was incubated at 4°C overnight with antibodies (6 µL anti-Zld, 8 µL anti-Grh, 10 µg anti-Twi, 7.5 µL anti-HA).
20 µL protein A beads (Dynabeads Protein A, ThermoFisher Scientific) blocked with BSA were added and samples were incubated at 4°C for 4 hours.
Beads were separated on a magnet and washed 3× with low salt wash buffer (10 mM Tris pH 7.6, 1 mM EDTA, 0.1% SDS, 0.1% Na-Deoxycholate, 1% Triton-X 100, 150 mM NaCl), 2× high salt wash buffer (10 mM Tris pH 7.6, 1 mM EDTA, 0.1% SDS, 0.1% sodium deoxycholate, 1% Triton-X 100, 300 mM NaC), 2× LiCl wash buffer (0.25 M LiCl, 0.5% NP40, 0.5% sodium deoxycholate), and 1× with TE buffer with NaCl (10 mM Tris-HCl pH 8.0, 1 mM EDTA, 50 mM NaCl).
Beads were resuspended in elution buffer (50 mM Tris-HCl pH 8.0, 1% SDS, 10 mM EDTA) and incubated at 65°C for 10 minutes.
The IP and input samples were treated with 4.5 µL RNAse A for 30 minutes.
5 µL Proteinase K was added and samples were incubated overnight at 65°C to reverse crosslinking.
DNA was isolated by phenol:chloroform extraction, precipitated with EtOH, and resuspended in 20 µL H~2~O.Preparation of sequencing libraries was performed using NEBNext Ultra II kit (NEB) with 7 PCR cycles for library amplification.
Sequencing was performed on the Illumina Hi-Seq4000 using 50bp single-end reads, or on the Illumina NovaSeq 6000 using 150bp paired-end reads.

## ChIP-seq analysis

Bowtie2 version 2.4.4 [@Langmead2012] was used to align ChIP-seq reads to the *Drosophila melanogaster* genome (version dm6) with the following non-default parameters: `-k 2 --very-sensitive --no-mixed --no-discordant -X 5000`.
Aligned reads were filtered to include only reads with a mapping quality score \> 30.
Reads mapping to unplaced scaffolds or the mitochondrial genome were discarded.
Peak calling was performed using MACS2 version 2.2 with parameters `-g dm --call-summits`.
Data from ChIP experiments performed with the same antibody in wild-type cells were used as a control for peak calling.
Thus, only peaks with statistically significantly enrichment (MACS2 q-value \< 0.5) in induced cells relative to wild-type cells are called as peaks.
Only peaks that were detected in all replicates were considered in downstream analysis.
bigWig files were generated using deepTools bamCoverage version 3.5.1 with parameters `–binSize 10`.
bigWig files were z-score normalized by subtracting the genome-wide average of all bins from each bin value and dividing by the standard deviation of all bins.

## RNA-seq

5×10^6^ cells were harvested and resuspended in 800 µL Trizol (Invitrogen).
RNA was purified using a chloroform extraction followed by a column-based Quick-RNA MiniPrep extraction (Zymo).
Poly-A-selected RNA sequencing (RNA-seq) libraries were prepared using the TruSeq RNA sample prep kit v2 (Illumina).

## RNA-seq analysis

Raw reads were aligned to the Drosophila melanogaster (dm6) genome using HISAT2 version 2.1.0 with parameters `-k 2 –very-sensitive`.
Reads with a mapping quality score \< 30 and reads aligning to the mitochondrial genome or unplaced scaffolds were discarded.
FeatureCounts (from Subread version 2.0) was used to assign aligned reads to genes (FlyBase gene annotation release r6.45).
The resultant table of read counts was used to perform differential expression analysis using DESeq2 [@love2014].
Genes with an adjusted p-value \< 0.05 and a fold change \> 2 were considered differentially expressed.

## ATAC-seq

ATAC-seq was performed as described previously [@Buenrostro2013].
2×10^5^ cells were harvested for ATAC-seq by centrifugation at 600 × *g* for 3 minutes at RT.
Cells were washed once in 100 µL 1X PBS and resuspended in 100 µL ATAC lysis buffer (10 mM Tris pH 7.5, 10 mM NaCl, 3 mM MgCl~2~,0.1% NP-40).
Cells were centrifuged at 600 × *g* for 10 minutes at 4°C.
Supernatant was removed and pellet was resuspended in 47.5 µL buffer TD (Illumina).
2.5 uL of Tagment DNA Enzyme (Illumina) was added and samples were incubated at 37°C for 30 minutes.
Tagmented DNA was purified using MinElute Reaction Cleanup Kit (Qiagen) and eluted in 10 µL of buffer EB.
DNA was PCR amplified for 12 cycles with the following conditions:72°C for 5 minutes, 98°C for 30 seconds, then 12 cycles of 98°C for 10 seconds, 63°C for 30 seconds and 72°C for 1 minute.
Amplified libraries were purified using a 1.2X ratio of Axygen paramagnetic beads.
Sequencing was performed on the Illumina NovaSeq 6000 using 150bp paired-end reads.

## ATAC-seq analysis

Raw ATAC-seq reads were trimmed to remove adapter sequences using NGmerge version 0.3 [@Gaspar2018].
Trimmed reads were aligned to the *Drosophila melanogaster* genome (version dm6) using bowtie2 with paramters `-k 2 --very-sensitive --no-mixed --no-discordant -X 5000`.
Reads with a mapping quality \< 30 and reads aligning to the mitochondrial genome or scaffolds were discarded.
As described previously, only fragments \< 100bp were considered for downstream analysis [@Buenrostro2013].
Reads were combined across all replicates and peak calling was performed using MACS2 with parameters `-f BAMPE --keep-dup all -g dm --call-summits`.
featureCounts was used to count the number of reads aligning within 200bp (100bp upstream or downstream) of peak summits.
Generation of z-score normalized bigWig files was performed as described above for ChIP-seq data.
Differential accessibility analysis was performed using DESeq2.
Regions with an adjusted p-value \< 0.05 were considered differentially accessible.

## Treatment of S2 cells with tazemetostat

Cells were plated at 1×10^6^ cells per mL and DMSO or tazemetostat (Fisher Scientific) was added to a final concentration of 10 µM.
Cells were split on days 2 and 5 and tazemetostat was added to maintain a final concentration of 10 µM.
On day 5, CuSO~4~ was added to achieve physiological expression of Zld, Grh or Twi (see above).
On day 7, cells were harvested for immunoblotting, ChIP and CUT&RUN.

## Drosophila strains and collection of brains for CUT&RUN

Drosophila strains and genetic crosses for the isolation of type II neuroblasts (neural stem cells) and for the ectopic expression of Zld in neural stem cells were described previously [@Larson2021a].
Brains from *brat^11/DF^* mutant larvae were used to perform CUT&RUN for endogenous Zld in neural stem cells.
For ectopic expression of Zld, brains were collected from *brat^11/DF,Tub-Gal80ts;UAS-HA-Zld^* larvae, in which *Worniu-Gal4* drives overexpression of HA-Zld in the neural stem cells.
Larvae were grown at 18°C until reaching the L3 stage, then shifted to 33°C for 24 hours to induce Zld overexpression.
50 brains were dissected into Schneider's medium.
Brains were centrifuged at 600 × *g* for 3 minutes and transferred to CUT&RUN wash buffer supplemented with 0.1% BSA.
Brains were dounced 10 times in a 1 mL dounce, centrifuged at 600 × *g* for 3 minutes and resuspended in 100 µL CUT&RUN wash buffer.

## CUT&RUN

CUT&RUN was performed using EpiCypher reagents according to the manufacturer protocol.
2×10^5^ cells, or material from 50 brains prepared as described above, were used for each CUT&RUN reaction.
Overnight incubation with antibodies was performed in antibody buffer containing 0.05% digitonin.
0.5 µL anti-H3K27me3 (Cell signaling technologies), anti-Zld, anti-Grh, or rabbit IgG were used for CUT&RUN.
For H3K27me3 CUT&RUN, a 1:5 dilution of K-MetStat Panel spike-in nucleosomes (EpiCypher, cat. \# 19-1002) was added to each reaction prior to addition of antibody.
Libraries were prepared using the NEBNext Ultra II kit (NEB).
During library preparation, cleanup steps were performed using a 1.1X ratio of Axygen paramagnetic beads, as recommended by EpiCypher.
PCR amplification was performed with the following conditions: 98°C for 45 seconds, then 14 cycles of 98°C for 15 seconds, 60°C for 10 seconds, and a final extension of 72°C for 1 minute.
Sequencing was performed on the Illumina NovaSeq 6000 with 150bp paired-end reads.

## CUT&RUN analysis

Trimming, alignment, and quality filtering of CUT&RUN reads was performed as described above for ATAC-seq data.
Peak calling was performed using MACS2 with parameters `–broad -f BAMPE –keep-dup all -g dm` for H3K27me3 or parameters `-f BAMPE --keep-dup all -g dm --call-summits` for Zld and Grh.
Generation of z-score normalized bigWig files was performed as described above for ChIP-seq data.
For H3K27me3 data, the percentage of reads containing barcoded Epicypher nucleosomes was calculated and 1 divided by this percentage was used as the scaling factor for spike-in normalization.
The z-score normalized read depth was multiplied by these scaling factors to generate spike-in normalized bigWigs.

## Motif analysis

Motif analysis was performed using the MEME suite[@Bailey2015], the memes R package[@Nystrom2021], and the universalmotif R package[@universalmotif].
To facilitate analysis of canonical Zld, Grh, or Twi motifs, position-weight matrices were derived from ChIP-seq data for each factor in the *Drosophila* embryos.
FIMO[@Grant2011] was used to find instances of the canonical motifs within ChIP-seq peaks for S2 cells.
Motifs with an adjusted p-value \< 1×10^-3^ were considered for analysis of motif number and score.
AME[@mcleay2010] was used to test for enrichment of known motifs from the FlyFactorSurvey database[@zhu2010].
Previously published RNA-seq data was used to pre-filter motifs to include only those for which the associate factor is expressed (RPKM \> 1) in at least one of the tissues being analyzed (embryo, neural stem cells, and wing discs).
Motifs enrichment was visualized using the normalized motif rank implemented by the memes package.

## Analysis of previously published data

Previously published datasets analyzed in this study can be found in Supplementary Table 2.
All previously published data were analyzed in parallel with data from this study using analysis parameters described above.

## Data and code availability

Sequencing data have been deposited in the Gene Expresion Omnibus under accession GSE227884.
Reviewers can access the data using reviewer token whsjesemxzqllyj.
All analysis code is available on GitHub: https://github.com/tjgibson/S2_pioneers_manuscript.

## Acknowledgements

We thank Alex Theis, Meghan Freund, and Andrew Mehle for helpful discussions and advice.
We thank Julia Zeitlinger for generously sharing the Twist antibody.
We particularly thank Elizabeth Larson for sharing the neural stem cell data for HA-Zld.
We acknowledge the University of Wisconsin-Madison Biotechnology Center and the NUSeq Core Facility for sequencing.
TJG was supported by the National Institutes of Health (NIH) National Research Service Award T32 GM007215.
Experiments were supported by a R35 GM136298 and R01 NS111647 from NIH to MMH.
MMH was also supported by a Vallee Scholar Award.
MMH is a Romnes Faculty Fellow and Vilas Faculty Mid-Career Investigator.